Contextual Combinatorial Cascading Bandits

### 2015-11-20 contextual combinatorial cascading bandits.tex
update basics for 'and' case

	At time $t$, the learning agent is revealed with feature vectors $X_{t,a} \in \RR^d$ for every basic arm $a \in E$, where we have $\norm{X_{t,a}}_2 \leq 1$; this feature vector combines both information of the user and the corresponding basic arm. We use Bernoulli random variable $\bw_{t}(a) \in \{0,1\}$, which is called weight for arm $a$ at time $t$, to indicate whether the user on round $t$ will click on the item $a$ or not. Assume $\bw_{t,a}$ are mutually independent and satisfy
	\begin{equation}
	\label{eq:expectation}
	\EE[\bw_{t,a} | X_{t,a}] = \theta_*^{\top} X_{t,a}
	\end{equation}
	where $\theta_*$ is an unknown $d$-dimensional vector with the assumption that $\norm{\theta_*}_2 \leq 1$ and $0 < \theta_*^{\top} x_{t,a} < 1$ for all $t, a$. At time $t$, the learning agent chooses a solution $\bA_t=(\ba_{1}^t,...,\ba_{\abs{\bA_t}}^t) \in \cS$ based on its past observations. The user then checks from the first item and stops if she has checked all items or click on one interesting. Suppose we add position discount $1 \geq\gamma_1 \geq \gamma_2\geq \cdots \geq \gamma_K>0$. If the user clicks on the $k$-th item of $\bA_t$, the reward we receive is $\gamma_k$. At the end of time $t$, the agent observes $\bO_t$ items in $\bA_t$ and receive the reward
	$$
	\br_t = \max_{1 \leq k \leq \abs{\bA_t}} \gamma_k \bw_t(\ba_k^t) = \bigvee_{k=1}^{\abs{A_t}} \gamma_k \bw_t(\ba_k^t),
	$$
	where we use the notation that $\bigvee_{1\leq i\leq n}a_i = \max_{1\leq i\leq n}a_i$. Note that every time $t$, we have observed $\bw_t(\ba_k^t), 1\leq k\leq\bO_t$.

	If we define a function $f$ on $A=(a_1,...,a_{\abs{A}}) \in \cS, w=(w(1),...,w(L))$ by
	\begin{align*}
	& f : \cS \times [0,1]^E \to [0,1]\\
	& f(A,w) = \sum_{k=1}^{\abs{A}}\gamma_k (\prod_{i=1}^{k-1}(1 -w(a_i)))w(a_k),
	\end{align*}
	then we have $\br_t = f(\bA_t, \bw_t)$ and $\EE[r_t]=f(\bA_t,\theta_*^{\top}X_t)$ where $X_t=(X_{t,1} \cdots X_{t,L})\in\RR^{d\times L}$. Let 
	$$
	A_t^* = \argmax_{A\in \cS} f(A,\theta_*^{\top}X_t).
	$$ 
	The goal of our learning algorithm is to minimize the expected cumulative regret in $T$ steps
	$$
	R(T) = \EE[\sum_{t=1}^T f(A_t^*, \theta_*^{\top}X_t) - f(\bA_t, \theta_*^{\top}X_t)].
	$$
	In the rest of this paper, we denote $\Delta_{t,\bA_t} = f(A_t^*, \theta_*^{\top}X_t) - f(\bA_t, \theta_*^{\top}X_t)$ and $R(T) = \EE[\sum_{t=1}^T \Delta_{t,\bA_t}]$.

@@

	\begin{lemma}
		\label{lem:increasing} 
		$1\geq \gamma_1\geq \cdots \geq \gamma_K > 0$. If we have $w \leq w'$ as column vectors in $\RR^L$, then for any $A\in\cS$,
		$$
		f(A, w) \leq f(A, w').
		$$
	\end{lemma}
	\begin{proof}
		Because for $1 \leq k \leq m$,
		\begin{align*}
		&~~\gamma_k - \sum_{i=k+1}^m \gamma_i\prod_{j=k+1}^{i-1}(1-w_j)w_i\\
		&\geq \gamma_k (1 - \sum_{i=k+1}^m \prod_{j=k+1}^{i-1}(1-w_j)w_i) \\
		&=\gamma_k \prod_{i=k+1}^{m}(1-w_i)\geq 0,
		\end{align*}
		then
		\begin{align*}
		&~~\gamma_k w_k + (1-w_k)\sum_{i=k+1}^m \gamma_i\prod_{j=k+1}^{i-1}(1-w_j)w_i\\
		& \leq \gamma_k w_k' + (1-w_k')\sum_{i=k+1}^m \gamma_i\prod_{j=k+1}^{i-1}(1-w_j)w_i.
		\end{align*}
		Therefore, 
		\begin{align*}
		& f(A; w_1',...,w_{k-1}',w_k,w_{k+1},...,w_m)\\
		&=\sum_{i=1}^{k-1} \gamma_i\prod_{j=1}^{i-1}(1-w_j')w_i' + \prod_{j=1}^{k-1}(1-w_j')[\gamma_k w_k \\
		&\qquad +(1-w_k)\sum_{i=k+1}^{m}\gamma_{i}\prod_{j=k+1}^{i-1}(1-w_j)w_i]\\
		&\leq \sum_{i=1}^{k-1} \gamma_i\prod_{j=1}^{i-1}(1-w_j')w_i' + \prod_{j=1}^{k-1}(1-w_j')[\gamma_k w_k' \\
		&\qquad +(1-w_k')\sum_{i=k+1}^{m}\gamma_{i}\prod_{j=k+1}^{i-1}(1-w_j)w_i]\\
		&=f(A; w_1',...,w_{k-1}',w_{k}',w_{k+1},...,w_m).
		\end{align*}
		Then we obtain the result.
	\end{proof}

@@

	\begin{lemma}
		With high probability $1-\delta$, $E[\Delta_{\bA_t}|\cH_t] \leq \frac{2}{f_t^*} \beta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma^{k-1}X_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}$
	\end{lemma}
	\begin{proof}
		Using $f(\bA_t^*,\bU_t) \leq f(\bA_t,\bU_t)$, then
		\begin{align*}
		&\sum_{k=1}^{\abs{A_t^*}}\gamma_k (\prod_{i=1}^{k-1}(1 -\bU_t(a_{t,i}^*)))\bU_t(a_{t,k}^*) \\
		\leq &\sum_{k=1}^{\abs{\bA_t}}\gamma_k (\prod_{i=1}^{k-1}(1 -\bU_t(\ba_i^t)))\bU_t(\ba_k^t) 
		\end{align*}
		By lemma \ref{lem:estimateU} and lemma \ref{lem:increasing}, it holds that
		\begin{align*}
		&~~f(A_t^*, \theta_*^{\top}X_t) \\
		&\leq f(\bA_t, \min\{\theta_*^{\top}X_{t,a}+2\beta_{t-1}(\delta)\norm{X_{t,a}}_{\bV_{t-1}^{-1}},1\}).
		\end{align*}
		Therefore by lemma \ref{lem:estimateTech},
		\begin{align*}
		&~~f(A_t^*, \theta_*^{\top}X_t) \\
		&\leq  \sum_{k=1}^{\abs{\bA_t}}\gamma_k (\prod_{i=1}^{k-1}(1 -\theta_*^{\top}X_{t,a_{t,i}^*})\theta_*^{\top}X_{t,a_{t,k}^*} \\
		&~~ + 2\sum_{k=1}^{\abs{\bA_t}}\gamma_k\beta_{t-1}(\delta)\norm{X_{t,a}}_{\bV_{t-1}^{-1}}\\
		&=f(\bA_t, \theta_*^{\top}X_t) + 2\beta_{t-1}(\delta)\sum_{k=1}^{\abs{\bA_t}}\gamma_k\norm{X_{t,a}}_{\bV_{t-1}^{-1}},
		\end{align*}
		which is equivalent to 
		$$
		\Delta_{\bA_t} \leq 2\sum_{k=1}^{\abs{\bA_t}}\gamma_k\beta_{t-1}(\delta)\norm{X_{t,a}}_{\bV_{t-1}^{-1}}.
		$$
	\end{proof}

@@

>>>>

At time $t$, the learning agent is revealed with feature vectors $X_{t,a} \in \RR^d$ for every basic arm $a \in E$, where we have $\norm{X_{t,a}}_2 \leq 1$; this feature vector combines both information of the user and the corresponding basic arm. We use Bernoulli random variable $\bw_{t}(a) \in \{0,1\}$, which is called weight for arm $a$ at time $t$, to indicate whether the user on round $t$ will click on the item $a$ or not. Assume $\bw_{t,a}$ are mutually independent and satisfy
\begin{equation}
\label{eq:expectation}
\EE[\bw_{t,a} | X_{t,a}] = \theta_*^{\top} X_{t,a}
\end{equation}
where $\theta_*$ is an unknown $d$-dimensional vector with the assumption that $\norm{\theta_*}_2 \leq 1$ and $0 < \theta_*^{\top} x_{t,a} < 1$ for all $t, a$. At time $t$, the learning agent chooses a solution $\bA_t=(\ba_{1}^t,...,\ba_{\abs{\bA_t}}^t) \in \cS$ based on its past observations. The user then checks from the first item and stops if she has checked all items or click on one interesting. Suppose we add position discount $0 \leq \gamma_1,...,\gamma_K\leq \gamma_0 \leq 1$. If the $k$-th item of $\bA_t$ appears blocking, the reward we receive is $\gamma_k$. And if the whole path of $\bA_t$ is unblocked, the received reward is $\gamma_0$. At the end of time $t$, the agent observes $\bO_t$ items in $\bA_t$ and receive the reward
\begin{align*}
\br_t &= \min_{1 \leq k \leq \abs{\bA_t}} [\gamma_k (1 - \bw_t(\ba_k^t)) +\gamma_0 \bw_t(\ba_k^t)] \\
&= \bigwedge_{k=1}^{\abs{A_t}} [\gamma_k (1 - \bw_t(\ba_k^t)) +\gamma_0 \bw_t(\ba_k^t)],
\end{align*}
where we use the notation that $\bigwedge_{1\leq i\leq n}a_i = \min_{1\leq i\leq n}a_i$. Note that every time $t$, we have observed $\bw_t(\ba_k^t), 1\leq k\leq\bO_t$.

If we define a function $f$ on $A=(a_1,...,a_{\abs{A}}) \in \cS, w=(w(1),...,w(L))$ by
\begin{align*}
f : \cS \times [0,1]^E \to [0,1]&~\\
f(A,w;\gamma_1,...,\gamma_K,\gamma_0) =& \sum_{k=1}^{\abs{A}}\gamma_k (\prod_{i=1}^{k-1}w(a_i))(1 - w(a_k)) \\
\qquad \qquad &~+ \gamma_0 \prod_{i=1}^{\abs{A}}w(a_i),
\end{align*}
then we have $\br_t = f(\bA_t, \bw_t)$ and $\EE[\br_t]=f(\bA_t,\theta_*^{\top}X_t)$ where $X_t=(X_{t,1} \cdots X_{t,L})\in\RR^{d\times L}$. Let 
$$
A_t^* = \argmax_{A\in \cS} f(A,\theta_*^{\top}X_t).
$$ 
The goal of our learning algorithm is to minimize the expected cumulative regret in $T$ steps
$$
R(T) = \EE[\sum_{t=1}^T f(A_t^*, \theta_*^{\top}X_t) - f(\bA_t, \theta_*^{\top}X_t)].
$$
In the rest of this paper, we denote $\Delta_{t,\bA_t} = f(A_t^*, \theta_*^{\top}X_t) - f(\bA_t, \theta_*^{\top}X_t)$ and $R(T) = \EE[\sum_{t=1}^T \Delta_{t,\bA_t}]$.

@@

\begin{lemma}
\label{lem:increasing} 
If $1 \leq \gamma_1 \leq ... \leq \gamma_K \leq \gamma_0 \leq 1$ and we have $w \leq w'$ as column vectors in $\RR^L$, then for any $A\in\cS$,
$$
	f(A, w) \leq f(A, w').
$$
\end{lemma}
\begin{proof}
	It suffices to prove that $\cA = [m]$ where $1 \leq m \leq K$. Because for $1 \leq k \leq m$,
	\begin{align*}
		&~~-\gamma_k + \sum_{i=k+1}^m \gamma_i(1- w_i')\prod_{j=k+1}^{i-1}w_j' + \gamma_0\prod_{j=k+1}^{m}w_j'\\
		&\geq \gamma_k (-1 + \sum_{i=k+1}^m (1- w_i')\prod_{j=k+1}^{i-1}w_j' +  \prod_{j=k+1}^{m}w_j')\\
		&=\gamma_k \cdot 0 = 0,
	\end{align*}
	then
	\begin{align*}
		&~~\gamma_k (1 - w_k) + w_k\sum_{i=k+1}^m \gamma_i(1- w_i')\prod_{j=k+1}^{i-1}w_j'\\
		&\qquad \qquad + \gamma_0 w_k\prod_{j=k+1}^{m}w_j'\\
		& \leq \gamma_k (1 - w_k') + w_k'\sum_{i=k+1}^m \gamma_i(1- w_i')\prod_{j=k+1}^{i-1}w_j'\\
		&\qquad \qquad + \gamma_0 w_k'\prod_{j=k+1}^{m}w_j'.
	\end{align*}
	Therefore, 
	\begin{align*}
		& f(A; w_1,...,w_{k-1},w_k,w_{k+1}',...,w_m')\\
		&=\sum_{i=1}^{k-1} \gamma_i(1 - w_i)\prod_{j=1}^{i-1}w_j + \prod_{j=1}^{k-1}w_j [\gamma_k (1 - w_k) \\
		&\qquad + w_k\sum_{i=k+1}^m \gamma_i(1- w_i')\prod_{j=k+1}^{i-1}w_j' + \gamma_0 w_k\prod_{j=k+1}^{m}w_j'\\
		&\leq \sum_{i=1}^{k-1} \gamma_i(1 - w_i)\prod_{j=1}^{i-1}w_j + \prod_{j=1}^{k-1}w_j [\gamma_k (1 - w_k') \\
		&\qquad + w_k'\sum_{i=k+1}^m \gamma_i(1- w_i')\prod_{j=k+1}^{i-1}w_j' + \gamma_0 w_k'\prod_{j=k+1}^{m}w_j'\\
		&=f(A; w_1,...,w_{k-1},w_{k}',w_{k+1}',...,w_m').
	\end{align*}
	Then we obtain the result.
\end{proof}

@@

Let $p_A$ denote the probability that all of items in $A$ are observed. Then $p_A = \prod_{k=1}^{\abs{A}-1}w(a_i)$.

\begin{lemma}

\end{lemma}

\begin{proposition}
	With high probability $1-\delta$, $E[\Delta_{\bA_t}|\cH_t] \leq \frac{2}{f_t^*} \beta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma^{k-1}X_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}$
\end{proposition}
\begin{proof}
	Using $f(\bA_t^*,\bU_t) \leq f(\bA_t,\bU_t)$, then
	\begin{align*}
		&\sum_{k=1}^{\abs{A_t^*}}\gamma_k (\prod_{i=1}^{k-1}(1 -\bU_t(a_{t,i}^*)))\bU_t(a_{t,k}^*) \\
		\leq &\sum_{k=1}^{\abs{\bA_t}}\gamma_k (\prod_{i=1}^{k-1}(1 -\bU_t(\ba_i^t)))\bU_t(\ba_k^t) 
	\end{align*}
		By lemma \ref{lem:estimateU} and lemma \ref{lem:increasing}, it holds that
		\begin{align*}
		&~~f(A_t^*, \theta_*^{\top}X_t) \\
		&\leq f(\bA_t, \min\{\theta_*^{\top}X_{t,a}+2\beta_{t-1}(\delta)\norm{X_{t,a}}_{\bV_{t-1}^{-1}},1\}).
		\end{align*}
		Therefore by lemma \ref{lem:estimateTech},
		\begin{align*}
		&~~f(A_t^*, \theta_*^{\top}X_t) \\
		&\leq  \sum_{k=1}^{\abs{\bA_t}}\gamma_k (\prod_{i=1}^{k-1}(1 -\theta_*^{\top}X_{t,a_{t,i}^*})\theta_*^{\top}X_{t,a_{t,k}^*} \\
		&~~ + 2\sum_{k=1}^{\abs{\bA_t}}\gamma_k\beta_{t-1}(\delta)\norm{X_{t,a}}_{\bV_{t-1}^{-1}}\\
		&=f(\bA_t, \theta_*^{\top}X_t) + 2\beta_{t-1}(\delta)\sum_{k=1}^{\abs{\bA_t}}\gamma_k\norm{X_{t,a}}_{\bV_{t-1}^{-1}},
		\end{align*}
		which is equivalent to 
		$$
		\Delta_{\bA_t} \leq 2\sum_{k=1}^{\abs{\bA_t}}\gamma_k\beta_{t-1}(\delta)\norm{X_{t,a}}_{\bV_{t-1}^{-1}}.
		$$
	\end{proof}

@@

\begin{lemma} % det(V_t)
	$\det(\bV_t) \leq (\lambda + C_\gamma t/d)^d.$
\end{lemma}
\begin{proof}
	We have $\det(\bV_t) \leq (\trace(\bV_t)/d)^d$ and
	\begin{align*}
	\trace(\bV_t) &= \trace(\lambda I) + \sum_{s=1}^t \sum_{k=1}^{\min\{\bO_s,\abs{\bA_s}\}} \gamma^{2k-2} \trace(X_{s,\ba_k^s} X_{s,\ba_k^s}^{\top})\\
	& = d \lambda + \sum_{s=1}^t \sum_{k=1}^{\min\{\bO_s,\abs{\bA_s}\}} \gamma^{2k-2} \norm{X_{s,\ba_k^s}}_2^2\\
	& \leq d \lambda + \sum_{s=1}^t\sum_{k=1}^{K}\gamma^{2k-2}=d\lambda + tC_\gamma.
	\end{align*}
\end{proof}

\begin{lemma} % technical lemma for det
	Let $x_i \in \RR^{d \times 1}, 1 \leq i \leq n$. Then we have
	$$
	\det(I + \sum_{i=1}^n x_i x_i^{\top}) \geq 1 + \sum_{i=1}^n \norm{x_i}_2^2.
	$$
\end{lemma}
\begin{proof}
	Denote the eigenvalues of $I + \sum_{i=1}^n x_i x_i^{\top}$ by $1+\alpha_1,...,1+\alpha_d$ with $\alpha_j \geq 0$, $1\leq j\leq d$. Then
	\begin{align*}
	\det(I + \sum_{i=1}^n x_i x_i^{\top}) &= \prod_{j=1}^d (1 + \alpha_j)\geq 1 +\sum_{j=1}^d \alpha_j =1-d + \sum_{i=1}^d (1+\alpha_i) \\
	&=1-d + \trace(I + \sum_{i=1}^n x_i x_i^{\top})= 1-d + d + \sum_{i=1}^n \norm{x_i}_2^2\\
	&=1 + \sum_{i=1}^n \norm{x_i}_2^2.
	\end{align*}
\end{proof}

\begin{lemma}
	If $\lambda \geq C_\gamma$, it holds that
	$$
	\sum_{s=1}^t \sum_{k=1}^{\min\{\bO_s, \abs{\bA_s}\}}\gamma^{2k-2}\norm{X_{s,\ba_{k}^s}}_{\bV_{s-1}^{-1}}^2 \leq 2\log(\frac{\det(\bV_t)}{\det(\lambda I)}).
	$$
\end{lemma}
\begin{proof}
	\begin{align*}
	\det(\bV_t) &= \det(\bV_{t-1} + \sum_{k=1}^{\min\{\bO_t, \abs{\bA_t}\}} \gamma^{2k-2}X_{t,\ba_k^{t}}X_{t, \ba_k^{t}}^{\top})\\
	&=\det(\bV_{t-1})\det(I + \sum_{k=1}^{\min\{\bO_t, \abs{\bA_t}\}} \gamma^{2k-2}\bV_{t-1}^{-1/2}X_{t,\ba_{k}^{t}} (\bV_{t-1}^{-1/2}X_{t,\ba_{k}^{t}})^{\top})\\
	&\geq \det(\bV_{t-1}) (1 + \sum_{k=1}^{\min\{\bO_t, \abs{\bA_t}\}} \norm{\gamma^{k-1}X_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}^2)\\
	&\geq \det(\lambda I)\prod_{s=1}^{t}(1 + \sum_{k=1}^{\min\{\bO_s, \abs{\bA_s}\}} \norm{\gamma^{k-1}X_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2)
	\end{align*}
	Using the fact that $ 2\log(1+u) \geq u$ for any $u \in [0,1]$ and 
	$$
	0 \leq \norm{\gamma^{k-1}X_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2 \leq \norm{\gamma^{k-1}X_{s,\ba_k^s}}_2^2/\lambda_{\min}(\bV_{s-1}) \leq \gamma^{2k-2}/\lambda,
	$$
	we get
	\begin{align*}
	\sum_{s=1}^t \sum_{k=1}^{\min\{\bO_s, \abs{\bA_s}\}}\norm{\gamma^{k-1}X_{s,\ba_{k}^s}}_{\bV_{s-1}^{-1}}^2 &\leq 2\sum_{s=1}^t\log(1 + \sum_{k=1}^{\min\{\bO_s, \abs{\bA_s}\}} \norm{X_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2)\\
	&\leq 2(\log(\det(\bV_t)) - \log(\det(\lambda I)))
	\end{align*}
\end{proof}
